---
title: "Using Bookdown to create a Security Now Ebook"
author: Mike Ching
date: "2017-07-10"
slug: securitynow-ebook
categories:
  - R
tags:
  - bookdown
  - securitynow
---

Recently I have been exploring the **blogdown** package to redo my website using [Hugo](https://gohugo.io), and it resurrected an old idea that I had. Security Now, one of my favorite podcasts, has transcriptions of each episode, and I thought it would be neat to put them into an e-book and read the old episodes that way. I tried using Calibre, but while it did work, I wasn't able to figure out how to format it nicely. The creator of **blogdown**, Yihui Xue, has also written a packages called **bookdown** which looked like it had the right features to make a proper looking e-book. 

```{r}
library(bookdown)
library(dplyr)
```

I downloaded the episode transcripts from GRC.com. 

## Obtaining Text
```
DIRECTORY <- "../datasets/sn501/"

# change the i to go from the first show number you want to the last show number. As of 1 July 2017, the most recent show is #618.

for(i in 500:599) {
  shortname <- paste0("sn-", i, ".txt")
  showname <- paste0("https://www.grc.com/sn/", shortname)
  print(shortname)
  download.file(showname, destfile = paste0(DIRECTORY, shortname), method = "curl")
}
```

I loaded an example file and put it into a [tibble.](https://cran.r-project.org/web/packages/tibble/vignettes/tibble.html)
```{r}
DIRECTORY <- "~/Dropbox/Mike/securitynow/shows/"
# list.files(DIRECTORY)
episode <- readLines(paste0(DIRECTORY, "sn-400.txt"))
episode[1:17]
```

From what I could tell, the first 13 lines appear to be formatted as metadata. These would be helpful for formatting as header text for bookdown to work on.

Bookdown requires the first line of the file to be a chapter title with a first level heading. I think a useful format would be the title. Other metadata like Series and Archive could be discarded since would probably be the same for every file.

Conceptually for each transcript, I had to discard lines 1-3 and 9, and move the title to the top. The rest would probably be ok to stay for now. But what if episodes were not identically formatted like this? I should use regular expressions.

```{r}
line_grc <- grep("^GIBSON", episode)
line_series <- grep("^SERIES", episode)
line_archive <- grep("^ARCHIVE", episode)

episode <- episode[-c(line_grc, line_series, line_archive)]

line_title <- grep("^TITLE", episode)
episode_length <- length(episode)
episode <- episode[c(line_title, 1:episode_length)]
episode <- episode[-(line_title + 1)]

episode <- gsub("\t", " ", episode)
episode <- sub("TITLE:  ", "# ", episode)
head(episode)
```

```{r}
writeLines(episode, paste0(DIRECTORY, "sn-400.Rmd"))
```

So that's the process for one episode, how 'bout 501-600?

```
for(i in 400:410) {
  episode <- readLines(paste0(DIRECTORY, "sn-", i, ".txt"))
  
  line_grc <- grep("^GIBSON", episode)
  line_series <- grep("^SERIES", episode)
  line_archive <- grep("^ARCHIVE", episode)

  episode <- episode[-c(line_grc, line_series, line_archive)]

  line_title <- grep("^TITLE", episode)
  episode_length <- length(episode)
  episode <- episode[c(line_title, 1:episode_length)]
  episode <- episode[-(line_title + 1)]

  episode <- gsub("\t", " ", episode)
  episode <- sub("TITLE:  ", "# ", episode)

  writeLines(episode, paste0(DIRECTORY, "sn-", i, ".Rmd"))
}
```

