---
title: 'VEX Scores and Google Sheets'
author: 'Michael Ching'
date: '2017-11-30'
slug: vex-scores
categories:
  - R
tags:
  - robotics
---

My son is in a robotics competition this year. He has to score points by driving a robot around a course and making it pick up and place things. I thought it might be nice to chart his progress.

The real learning issue for today is using the [`googlesheets` package](https://github.com/jennybc/googlesheets) to work with the data. I often enter data using this site and it would save me a lot of time to get the data directly from Google Sheets rather than downloading it as a csv. The data loading section of this is based on [this vignette](https://rawgit.com/jennybc/googlesheets/master/vignettes/basic-usage.html).
```{r}
suppressPackageStartupMessages(library("dplyr"))
library(ggplot2)
library(googlesheets)
suppressPackageStartupMessages(library(lubridate))
```
First I had to establish a connection with the Google account I use. The following commands brought up a browser window and I authenticated there. Once authenticated I was able to see a list of my Google Sheets, as if I had gone to the website.
```{r}
my_sheets <- gs_ls() # Establish connection with my Google account
my_sheets %>% glimpse() # View the resulting data
```

To select a sheet I had to get some attributes about it. I used the title to find it, but I could have used its key or even the URL.
```{r}
vex <- gs_title("VEX_practice_scores") # Create GS object 
vex
```

Once I had the right sheet, I was able to read it in directly into R. Magic!
```{r}
vex_scores <- gs_read(vex)
vex_scores
```

From there it was a straightforward process to visualize the team's performance. I created a plot of their performance over time. Sadly they don't seem to be getting any better. A solid score would be 105-125. They average in the 70s.

```{r}
vex_scores$Points

summary(vex_scores$Points)

ggplot(dat = vex_scores, aes(x = Date, y = Points, group = Date)) + 
  geom_boxplot() +
  ggtitle("Team VEX Scores over Time")
```

I really should make a quality control run chart but that's a job for another day.

## Conclusion
The `googlesheets` package is a easy way to get info out of Google Sheets into R. Good luck to the boy on Saturday, but it looks like he has his work cut out for him.