---
title: Runkeeper Data
author: ~
date: '2018-06-03'
slug: runkeeper-data
categories: 
  - R
tags: 
  - exercise
subtitle: ''
---

Last summer I started running again after about 8 years off. I had run from my sophomore year of high school until the fall of 2008 when it seemed like I had a meniscus injury and I stopped running. That was 17 years of running during which the longest break I had taken was maybe a few months. I always missed it, so when my son started training for a Scouting event, I decided I would keep him company. Lo and behold my knee felt ok, so I figured I would keep going. I've been using the Runkeeper app to track my runs and decided to look at what the data showed.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r}
dat <- tbl_df(read.csv("../datasets/cardioActivities.csv"))
dat
```

```{r}
dat$Date <- ymd_hms(dat$Date)
```

## Activities
I don't consistently tracka c
```{r}
table(dat$Type)
```

## Weekly Mileage
```{r}
dat %>%  
#  filter(Date < ymd(20180601)) %>% 
  filter(Type == "Running") %>%
  mutate(Week_obs = week(Date), Year_obs = year(Date)) %>%
  group_by(Year_obs, Week_obs) %>%
  summarize(Mileage = sum(Distance..mi.)) %>%
  ggplot(aes(x = Week_obs, y = Mileage)) + 
  facet_grid( ~ factor(Year_obs)) +
  geom_point()


```

