---
title: Annual Leadership Forum Voting
author: ~
date: '2018-03-18'
slug: alf-voting
categories: 
  - R
tags: 
  - AAP
  - voting
draft: TRUE
---

It's been a while since I posted, but this weekend I returned to the American Academy of Pediatrics Annual Leadership Forum and I was able to cite my analysis on the top 10 resolutions. 

In the spirit of looking further at the ALF data I decided to look at the voting records from this year. It seemed that most votes pass by at least 80% if not more. This includes the amendment votes.

## Methods
For each resolution or amendment, I recorded the number voting and the percentage voting in favor. Here's the exploratory analysis.

```{r}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
dat <- read.csv("../datasets/ALF_votes.csv")
dat <- tbl_df(dat)
dat
```

```{r}
dat %>% mutate(Yes_p = Yes/Total) %>% 
  ggplot(aes(x = Yes_p, color = Notes)) + geom_histogram(binwidth = 0.05)
```

It looks like most 

One hypothesis is that resolutions that have to be amended are less likely to pass than resolutions that do not require amendment. 

There were `r dat %>% filter(Consent == 0) %>% nrow()` votes taken on 78 resolutions and 4 late resolutions. The number of votes is less than the number of resolutions because `r dat %>% filter(Consent == 1) %>% nrow()` were placed on the consent calendar. 

Of all the resolutions, `r dat %>% filter(Consent == 1) %>% nrow()/82` were on the consent calendar.

Overall success rate for amendments
```{r}
dat %>% filter(Notes == "Amendment") %>%
  summarize(Amendment_success_rate = mean(Success, na.rm = T))
```

```{r}
dat %>% filter(Notes == "Resolution") %>%
  summarize(Resolution_success_rate = mean(Success, na.rm = T))
```

Success rate if not on consent calendar
```{r}
dat %>% filter(Notes == "Resolution" & Consent == 0) %>%
  summarize(Resolution_success_rate = mean(Success, na.rm = T))
```

Distribution of votes overall
```{r}
dat %>% filter(Consent == 0) %>%
  group_by(Notes) %>% 
  summarize(Count = n())

```


Distribution of votes for amendments
```{r}
dat %>% filter(Notes != "CC")
```


Distribution of votes for resolutions