---
title: Runkeeper Data
author: ~
date: '2018-06-03'
slug: runkeeper-data
categories: 
  - R
tags: 
  - exercise
subtitle: ''
---

Last summer I started running again after about 8 years off. I had run from my sophomore year of high school until the fall of 2008 when it seemed like I had a meniscus injury and I stopped running. I had run for 17 years, during which the longest break I had taken was maybe a few months. After I stopped, I always missed it, and when my son started training for a Scouting running achievement, I decided I would keep him company. Lo and behold my knee felt ok, so I figured I would keep going. I've been using the [Runkeeper](https://runkeeper.com/index) app to track my runs and decided to look at what the data showed.

```{r}
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
suppressPackageStartupMessages(library(lubridate))
```

I downloaded the data from the [Runkeeper website.](https://runkeeper.com/index) It came as a zip file containing all my GPS tracks plus a processed csv file.
```{r}
dat <- tbl_df(read.csv("../datasets/cardioActivities.csv"))
glimpse(dat)
```

```{r}
dat$Date <- ymd_hms(dat$Date)
```

## Weekly Mileage
There was a lot of data there but I thought the first thing to start with would be my weekly mileage. I wanted to see how that changed over the last year. 
```{r}
dat %>%  
#  filter(Date < ymd(20180601)) %>% 
  filter(Type == "Running") %>%
  mutate(Week_obs = week(Date), Year_obs = year(Date)) %>%
  group_by(Year_obs, Week_obs) %>%
  summarize(Mileage = sum(Distance..mi.)) %>%
  ggplot(aes(x = Week_obs, y = Mileage)) + 
  facet_grid( ~ factor(Year_obs)) +
  geom_point() +
  labs(title = "Weekly Mileage",
       x = "Week", 
       y = "Miles")
```

## Monthly Mileage
In [What I Talk about When I Talk about Running](https://www.amazon.com/dp/B0015DWJ8W/ref=dp-kindle-redirect?_encoding=UTF8&btkr=1) by Haruki Murakami, he talks about his monthly mileage, averaging 136-186 miles per month. My monthly averages weren't as impressive as his.

```{r}
dat %>%  
  filter(Date < ymd("20180601")) %>%
  filter(Type == "Running") %>%
  mutate(Month_obs = month(Date), Year_obs = year(Date)) %>%
  group_by(Year_obs, Month_obs) %>%
  summarize(Mileage = sum(Distance..mi.)) %>%
  ggplot(aes(x = Month_obs, y = Mileage)) + 
  facet_grid( ~ factor(Year_obs)) +
  geom_point() +
  labs(title = "Monthly Mileage",
       x = "Month", 
       y = "Miles")
```

## Calories Burned
I tracked my weekly calories burned, expecting that these would be highly correlated with mileage.

```{r}
dat %>%  
  filter(Type == "Running") %>%
  mutate(Week_obs = week(Date), Year_obs = year(Date)) %>%
  group_by(Year_obs, Week_obs) %>%
  summarize(Calories = sum(Calories.Burned)) %>%
  ggplot(aes(x = Week_obs, y = Calories)) + 
  facet_grid( ~ factor(Year_obs)) +
  geom_point() +
  labs(title = "Weekly Calories",
       x = "Week", 
       y = "Total Calories")
```

## Correlation between Calories and Miles
The app estimated calories per run. There was a linear correlation between calories and miles run, as displayed below.

```{r}
dat %>% filter(Type == "Running") %>%
  ggplot(aes(x = Distance..mi., y = Calories.Burned)) + geom_point() + geom_smooth(method = "lm") +
  labs(title = "Correlation between Calories and Miles Run",
       x = "Miles",
       y = "Calories")
```

This was a very linear relationship, even at the very extremes of the distances run. I fit a linear model and determined that a rough estimate of the energy expenditure rate was 136 calories per mile.

```{r}
x <- filter(dat, Type == "Running")
model1 <- lm(x$Calories.Burned ~ x$Distance..mi.)
summary(model1)
```

