---
title: Weight Training Data
author: ~
date: '2018-06-03'
slug: weight-training-data
categories: 
  - R
tags: 
  - exercise
  - weights
subtitle: ''
---

I lifted weights regularly for two years but stopped recently because I decided to run more instead. I tracked my weight lifting using an app called Strong, and since it seemed that I had reached my peak of weight lifting I figured I would look back on my exercise over this time.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r}
dat <- read.csv("../datasets/strong.csv")
dat <- tbl_df(dat)
dat
```

From a high level look at the data, I have a table with each row being a set of an exercise. It's pretty amazing to think that I did 4334 sets of various exercises over these two years.

# Data Cleaning
Looking at the data there were some basic cleaning issues. One was that the Date format was read in as a factor and I needed to change it into a format that can be manipulated. The other was that there were some variables that weren't needed like the workout name (corresponding to a preset workout), mi (miles?), and Seconds. I didn't use these often or consistently so there's not much to be gained there.
```{r}
dat$Date <- ymd_hms(dat$Date)
dat <- dat %>% select(Date, Exercise.Name, Set.Order, lb, Reps) 
```

Now I was ready to answer some basic questions.

## How many days did I lift and how long does the data cover
```{r}
max(dat$Date) - min(dat$Date)
```

```{r}
length(unique(dat$Date))
```

I completed 218 workouts over the span of 816 days. Recently I have been lifting very rarely.

```{r}
dat %>%  
  mutate(Month_obs = month(Date), Year_obs = year(Date)) %>%
  group_by(Year_obs, Month_obs) %>%
  summarize(sets = n()) %>%
  mutate(Yr_mo = as.Date(paste0(Year_obs, "-", Month_obs, "-01"))) %>% 
  ggplot(aes(x = Yr_mo, y = sets)) + geom_point() + 
  geom_smooth(method = "loess") +
  ggtitle("Number of Sets per Month")
  
month(dat$Date[1])
```


## What exercises did I do
```{r}
dat %>% group_by(Exercise.Name) %>% summarize(sets = n()) %>%
  arrange(desc(sets))
```

That seemed about right. I did about 3 sets usually, and I did some exercises just about every time (Lat Pulldown for example).

## What kind of reps was I doing per set
I examined the number of reps I did per set.
```{r}
summary(dat$Reps)
ggplot(dat, aes(x = Reps)) + geom_histogram(binwidth = 2)
```
There is a second peak at 50 from crunches, but I was surprised that I did one set of 100. What was that all about? Looks like I haven't been doing much crunches lately.

```{r}
dat %>% filter(Reps > 50) %>% arrange(desc(Reps))
```

Taking out the crunches, most of my sets were 8-12. I also needed to take out Planks and Decline Crunches

```{r}
dat %>% filter(Exercise.Name != "Crunch") %>% 
  filter(Exercise.Name != "Plank (Custom)") %>%
  filter(Exercise.Name != "Decline Crunch") %>%
  ggplot(aes(x = Reps)) + geom_histogram(binwidth = 2)
```

## How did the weight change over time
I got stronger then not so strong as the workouts spaced out. How to show this? One way was to look at an exercise over time. Bench press machine was a good one since I did this over a long period of time.

```{r}
dat %>% filter(Exercise.Name == "Bench Press (Machine)") %>%
  group_by(Date) %>% summarize(maxwt = max(lb)) %>%
  ggplot(aes(x = Date, y = maxwt)) + geom_point() + geom_smooth()
```

Maybe that's not a great example. 

```{r}
dat %>% filter(Exercise.Name == "Bicep Curl (Dumbbell)") %>%
  group_by(Date) %>% summarize(maxwt = max(lb)) %>%
  ggplot(aes(x = Date, y = maxwt)) + geom_point() + geom_smooth()

```
