---
title: Prime Rib Temperature
author: ~
date: '2019-12-24'
slug: prime-rib-temperature
categories: 
  - R
tags: 
  - cooking
  - primerib
type: ''
subtitle: ''
image: ''
---

We started a new family Christmas Eve tradition with roast prime rib. This year was a 6.61 lb bone on prime rib. 

```{r}
library(dplyr)
library(ggplot2)
dat <- tbl_df(read.csv("../datasets/Prime_rib_temp.csv"))
```

I followed the [Serious Eats recipe](https://www.seriouseats.com/2014/12/food-lab-guide-to-prime-rib.html).

I cut off the bone then salted and let the bones and meat dry brine overnight. I tied them back together around noon and put them in a 200 degree Fahrenheit oven. The temperature didn't budge for most of the first hour and then it was a reasonably linear rise in temperature until I took it out when the meat was 125 degrees. It went up to 135 while resting for an hour.

```{r}
ggplot(dat, aes(x = minutes, y = temperature)) + geom_point()
```

What a beautiful line! Taking off the first few observations, we can put on a regression line.

```{r}
dat %>% filter(temperature >45) %>% 
  ggplot(aes(x = minutes, y = temperature)) + 
  geom_point() + geom_smooth(method = "lm")
  
```

I decided to fit a regression model to that line to see how linear it was.

```{r}
dat2 <- dat %>% filter(temperature > 45)
model1 <- lm(temperature ~ minutes, dat2)
summary(model1)
```

The coefficient was 0.46 degrees per minute with an R squared of 0.999! 