---
title: Tenderfoot Paces
author: ~
date: '2019-02-16'
slug: tenderfoot-paces
categories: 
  - R
tags: 
  - scouting
subtitle: ''
---

I had to organize the Tenderfoot scouts in my troop to learn some orienteering. The first step was to calculate how long their paces are. We went to the track and measured their pace count per 100 meters at least twice per scout. I wondered what the distribution looked like.

I used dplyr method to group and summarize the data. It's a little overkill but it does make it super easy. It looks like the median length of a pace is just about 151 cm, with a mean of pretty much the same thing. 

```{r}
library(googlesheets)
suppressPackageStartupMessages(library(dplyr))
library(ggplot2)
```

```{r, cache = F}
# Change cache flag to F if updating data
my_sheets <- gs_ls()
fb <- gs_title("BoysPaces")
x <- gs_read(fb)
```

Here's what the raw data looked like.
```{r}
x <- as_tibble(x)
x
nboys <- length(unique(x$Name))
```

Here I calculated the mean paces per 100 meters per boy and how long their paces were in meters and feet.
```{r}
x %>% group_by(Name) %>%
  summarize(mean_paces_per_100m = mean(paces_per_100m)) %>%
  mutate(paces_m = 100/mean_paces_per_100m, 
         paces_ft = 328.084/mean_paces_per_100m)
```

To look at the distribution I plotted a boxplot.
```{r}
x %>% group_by(Name) %>%
  summarize(mean_paces_per_100m = mean(paces_per_100m)) %>%
  mutate(paces_m = 100/mean_paces_per_100m) %>%
  ggplot(aes(y = paces_m)) + geom_boxplot()
```

Here's the median and mean pace length for the boys.
```{r}
x %>% group_by(Name) %>%
  summarize(mean_paces_per_100m = mean(paces_per_100m)) %>%
  mutate(pace_length_m = 100/mean_paces_per_100m) %>%
  summarize(median_pace_length = median(pace_length_m))

x %>% group_by(Name) %>%
  summarize(mean_paces_per_100m = mean(paces_per_100m)) %>%
  mutate(pace_length_m = 100/mean_paces_per_100m) %>%
  summarize(mean_pace_length = mean(pace_length_m), sem_pace_length = sd(pace_length_m)/sqrt(nboys))
```

