---
title: Frog Bog Metrics
author: ~
date: '2019-04-22'
slug: frog-bog-metrics
categories: 
  - R
tags: 
  - games
type: ''
subtitle: ''
image: ''
---

The school carnival was a couple weekends ago, and I volunteered at a game called Frog Bog. Players use a hammer to hit a lever that launches a frog at a rotating set of lily pads. You score by getting a frog in the lily pad. Players get four frogs per turn. I thought it might be nice to know the probability of success to help understand how to distribute prizes.

Here's a video from a different carnival. We only have one rotating set of lily pads at our carnival.

<iframe width="560" height="315" src="https://www.youtube.com/embed/nDRnu2wQCbA" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Loading the data...
```{r}
library(googlesheets)
suppressPackageStartupMessages(library(dplyr))
library(binom)
```

```{r}
my_sheets <- gs_ls()
fb <- gs_title("Frog_bog")
x <- gs_read(fb)
```

Here is what the raw data look like.
```{r}
x <- tbl_df(x)
x
```

Summarized in table form we have the number of times each outcome happened per game of 4 tries.
```{r}
table(x)
```

To find the probability of success, we make a <b>big</b> assumption that success is completely random and independent of each player's skill. The probability of a success on any one attempt is thus successes divided by attempts.
```{r}
attempts <- nrow(x) * 4
successes <- sum(x$Successes)
psuccess <- successes/attempts
```

Below I calculate the probability of 4, 3, 2, 1, and 0 successes per game.
```{r}
data.frame(successes = 0:4, p = dbinom(0:4, size = 4, prob = psuccess))
```

If we want to take about 90% of the entry fee (return 10% to winners), we need to determine what each prize should be. If we have no prize for 0 frogs, you already are collecting `r round(100 * (1-psuccess)^4, 2)`% of the proceeds.

Let's say we have a small prize for 1 success, valued at a small percentage of the entry fee. We would have a medium prize for 2 successes and a big prize for 3 successes. The grand prize could be much larger than the entry fee since it only happens about 1 in 5800 tries. I set the absolute dollar returns in the vector below.

```{r}
fractions <- c(0, 0.1, 1, 3, 10)
```

Here's the expected payout, broken down by probability of successes in one 4 frog game.
```{r}
dbinom(0:4, size = 4, prob = psuccess) * fractions
```

Overall expected payout per game would be then:
```{r}
sum(dbinom(0:4, size = 4, prob = psuccess) * fractions)
```

That gets pretty close to the goal amount of proceeds. The actual entry fee was $2.50.

At a conservative estimate of one game per minute, over 10 hours per day, in a 2 day fair, that's `r 10*60*2` games. The game thus makes approximately $`r round(10*60*2*(2.50 - sum(dbinom(0:4, size = 4, prob = psuccess) * fractions)), 2)` or $`r round(60*(2.5-sum(dbinom(0:4, size = 4, prob = psuccess) * fractions)), 2)` per hour.