---
title: Frog Bog Metrics
author: ~
date: '2019-04-22'
slug: frog-bog-metrics
categories: 
  - R
tags: 
  - games
type: ''
subtitle: ''
image: ''
---

The school fair was a couple weekends ago, and I volunteered at a game called Frog Bog. Players use a hammer to hit a lever that launches a frog at a rotating set of lily pads. You score by getting a frog in the lily pad. Players get four frogs per turn. I thought it might be nice to know the probability of success to help understand how to distribute prizes.

<iframe width="560" height="315" src="https://www.youtube.com/embed/nDRnu2wQCbA" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Loading the data...
```{r}
library(googlesheets)
suppressPackageStartupMessages(library(dplyr))
library(binom)
```

```{r}
my_sheets <- gs_ls()
fb <- gs_title("Frog_bog")
x <- gs_read(fb)
```

Here is what the raw data look like.
```{r}
x <- tbl_df(x)
x
```

Summarized in table form we have the number of times each outcome happened per game of 4 tries.
```{r}
table(x)
```

To find the probability of success, we make a big assumption that success is completely random and independent of each player's skill. The probability of a success on any one attempt is thus successes divided by attempts.
```{r}
attempts <- nrow(x) * 4
successes <- sum(x$Successes)
psuccess <- successes/attempts
```

Below I calculate the probability of 4, 3, 2, 1, and 0 successes per game.
```{r}
data.frame(successes = 0:4, p = dbinom(0:4, size = 4, prob = psuccess))
```

If we want to take about 80% of the proceeeds, we need to determine what each prize should be. Let's say the entry fee is $1. If we have no prize for 0 frogs, you already are collecting `r (1-psuccess)^4` of the proceeds.

Let's say we have a small prize for 1 success, valued at a small percentage of the entry fee. A medium prize for 2 successes, and a big prize for 3 successes. The grand prize could be much larger than the entry fee. I set the percentage of returns in the vector below.
```{r}
fractions <- c(0, 0.1, 1, 4, 10)
```

Here's the amount won for each outcome.
```{r}
dbinom(0:4, size = 4, prob = psuccess) * fractions
```

Overall expected winnings would be then:
```{r}
sum(dbinom(0:4, size = 4, prob = psuccess) * fractions)
```

