---
title: Restarting with California Health Interview Survey
author: ~
date: '2020-01-13'
slug: restarting-with-california-health-interview-survey
categories: 
  - R
tags: 
  - health
type: ''
subtitle: ''
image: ''
---

Lately this blog has had a lot of descriptive posts about my life but not much that stretches my R abilities. Later this year my organization is holding its research day, so I decided to start practicing on a dataset that I last used about 7 years ago. The [California Health Interview Survey](https://healthpolicy.ucla.edu/Pages/home.aspx) is a survey of households that is weighted to be representative of the non-institutionalized population of California. This dataset has a few wrinkles in it that require a higher level of R programming so I thought I would try to stretch my rusty skills a little.

```{r}
library(dplyr)
library(survey)
library(foreign)
```

Fortunately when I last used this dataset in 2013, I had commented my code pretty well so I was able to remember how to load in the Stata file.

```{r}
chis <- read.dta("/Users/mching/Dropbox/Mike/CHIS18/CHILD.dta", 
                 convert.dates = TRUE, convert.factors = TRUE, 
                 missing.type = TRUE, convert.underscore = TRUE, 
                 warn.missing.labels = FALSE)
```

# Unweighted Data

Here's what the data dictionary said about a variable documenting children who needed dental care in the last year. 

<table style="width:100%">
  <tr>
    <th> </th>
    <th> </th> 
    <th>Frequency</th>
    <th>%</th>
  </tr>
  <tr>
    <td>-1</td>
    <td>Inapplicable</td>
    <td>58</td>
    <td>3.66</td>
  </tr>
  <tr>
    <td>1</td>
    <td>Yes</td>
    <td>67</td>
    <td>4.22</td>
  </tr>
  <tr>
    <td>2</td>
    <td>No</td>
    <td>1461</td>
    <td>92.12</td>
  </tr>
  
</table>
  
And here's what the dataset I loaded had to say about this variable.
  
```{r}
table(chis$cb27)
round(prop.table(table(chis$cb27))*100, 2)
```

As you can see from above, the unweighted data were the same as in the data dictionary.

# Weighted Data

In the survey package you have to create a survey object that can be used to run the survey methods. Code for creating the survey object was nodified from AJ Damico's [Analyze Survey Data for Free](http://asdfree.com/california-health-interview-survey-chis.html) website.

```{r}
chis_design <- 
    svrepdesign( 
        data = chis , 
        weights = ~ rakedw0 , 
        repweights = "rakedw[1-9]" , 
        type = "other" , 
        scale = 1 , 
        rscales = 1 , 
        mse = TRUE 
    )
```

I generated a quick table to see if the survey method for calculating the population estimates worked. 

```{r}
round(svytable(~cb27, chis_design), -3)
```

I checked it against [AskCHIS](http://ask.chis.ucla.edu/), the web interface to the CHIS dataset. Here's what it returned for this variable: 

<table style="width:100%">
  <tr>
    <th>Needed Dental Care</th>
    <th>Percentage</th> 
    <th>Confidence Interval (95%)</th>
    <th>Population</th>
  </tr>
  <tr>
    <td>Needed dental care</td>
    <td>4.3%</td>
    <td>2.5 - 6.2</td>
    <td>247,000</td>
  </tr>
  <tr>
    <td>Did not need dental care</td>
    <td>95.7%</td>
    <td>93.8 - 97.5</td>
    <td>5,477,000</td>
  </tr>
  <tr>
    <td>Total</td>
    <td>100.0%</td>
    <td>-</td>
    <td>5,724,000</td>
  </tr>
  
</table>

The estimated population  number of children who needed/didn't neet care was exactly the same as what I calculated. The ratios were not the same because of all the inapplicable results.

```{r}
svymean(~cb27, chis_design)
```

Getting the weighted frequencies on just the yes or no responses was tricky. It required creating a variable for the Inapplicable patients and then creating a mean based on that.

```{r}
chis$cb27_inapplicable <- ifelse(chis$cb27 == "INAPPLICABLE", T, F)

chis_design <- 
    svrepdesign( 
        data = chis , 
        weights = ~ rakedw0 , 
        repweights = "rakedw[1-9]" , 
        type = "other" , 
        scale = 1 , 
        rscales = 1 , 
        mse = TRUE 
    )

svyby(~cb27, ~cb27_inapplicable, svymean, design = chis_design)
```

The weighted means were the same as what AskCHIS said (yes = 4.3%, no = 95.7%). The confidence intervals were very slightly off from AskCHIS (yes = 2.5-6.2, no = 93.8-97.5) but I figured they were close enough to not matter too much.

```{r}
confint(svyby(~cb27, ~cb27_inapplicable, svymean, design = chis_design))
```

# Discussion
I was able to load the CHIS 2018 child dataset and confirm that I could use the survey methods to replicate the results from the official CHIS web interface. I didn't do too much with the data, but next time I will pick up with doing some plots and regressions.