---
title: Finisher Characteristics and Hawaii Update
author: ~
date: '2020-05-30'
slug: finisher-characteristics-and-hawaii-update
categories: []
tags: 
  - exercise
  - running
type: ''
subtitle: ''
image: ''
---

I'm almost at the end of May, and I'm over 100 miles into this race. It's been a bit tough so far because of the time trial I did last week, but I'm still trying to get some cushion on the August 31 deadline for the race.

On the Facebook group for this race they talk about the races between races. Some people look at their state standings. Others may do their name rankings (i.e., rank among people named Smith or Mike).

I decided to look at how I've been doing compared to the other Hawaii runners and the runners with my first name. I also wanted to look at the people who have already finished and how they compare to the people who are still running.

# Data Loading and Cleaning
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(epiR))
```

```{r}
gv <- read_csv("../datasets/gvrat_20200530.csv") # Active runners in GVRAT
bv <- read_csv("../datasets/gvbat_20200530.csv") # Finshers who are going back

bv <- bv %>% mutate(finish = T)
gv <- gv %>% mutate(finish = F)


# Combine the finisher table and the active participant tables
gv <- bind_rows(gv, bv)

# I got rid of the doggies running in the doggy virtual race
gv <- gv %>% filter(Event == "GVRAT")


# Make data into long form for dplyr use
gv_miles <- gv %>% 
  select(-c(Position, `Participant's Name`, `Your Approximate   Location`,
            `Comp%`, `Proj Fin`, KM, Home, G, A, Miles)) %>% 
  pivot_longer(contains("/"), names_to = "run_date", values_to = "miles_d") %>%
  mutate(run_date = mdy(paste0(run_date, "/2020"))) %>%
  mutate(Event = as.factor(Event))

gv_roster <- gv %>% 
  select(Bib, name = `Participant's Name`, Event, G, A, Home, Miles, finish) %>%
  mutate(Event = as.factor(Event)) %>%
  mutate(Home = as.factor(Home)) %>%
  mutate(G = as.factor(G)) 

rm(bv)
rm(gv)
```

# Hawaii Runners

There are now 42 runners from Hawaii.

```{r}
Hawaii <- gv_roster %>% filter(Home == "USA - Hawaii") %>% 
  mutate(place = dense_rank(-Miles)) %>%
  select(Bib, name, G, A, Miles, place)

Hawaii
```

Today, I'm in place #`r Hawaii$place[which(Hawaii$name == "Michael Ching")]` among Hawaii runners. 

These Hawaii people seem to be doing pretty good. No one is going crazy, with the clear leader having about 100 miles per week.

I decided to look at my rank compared to the other runners over time.

```{r}
gv_miles %>% filter(Bib %in% Hawaii$Bib) %>% 
  arrange(run_date) %>% 
  group_by(Bib) %>%
  mutate(miles_t = cumsum(miles_d)) %>%
  ungroup %>%
  group_by(run_date) %>%
  mutate(place_d = min_rank(-miles_t)) %>%
  filter(Bib == 18986) %>%
  ggplot(aes(x = run_date, y = place_d)) + geom_line() +
  ylab("My place among Hawaii Runners") + xlab("Date")
```

# The Michaels

I could do the same thing for my name "Michael".
```{r}
Michael <- gv_roster %>% filter(grepl("Michael", name)) %>% 
  mutate(place = dense_rank(-Miles)) %>%
  select(Bib, name, G, A, Miles, place)

Michael
```

There are `r nrow(Michael)` Michaels in the race! Thanks in part to my late start I am in `r Michael$place[which(Michael$Bib == 18986)]` place among Michaels.

Here are some other Michael characteristics.

Michael ages. Not too many young Michaels anymore since Michael hasn't been as popular lately.

```{r}
Michael %>% ggplot(aes(x = A)) +
  geom_histogram(binwidth = 5)

summary(Michael$A)
```

Michaels are doing great with their miles. If we were one giant commingled Michael who is running at the mean pace, we would be on track to finish the race.
```{r}
summary(Michael$Miles)
```

# Finisher Characteristics

## Age and Gender for Finishers:
```{r}
gv_roster %>% group_by(finish) %>%
  summarize(n(), mean_age = mean(A, na.rm = T), 
            median_age = median(A, na.rm = T),
            male = mean(as.numeric(G)-1))
```
So far, finishers are older than non-finishers and are more likely to be male.

## Bivariate Comparisons

The 95% confidence interval of the difference in age is -6.9 to -0.8 years.

```{r}
t.test(gv_roster$A ~ gv_roster$finish)
```

Now looking at the effect that gender has on finishing, we can see that females are about 1/3 as likely as males to have finished the race.

```{r}
tab1 <- table(gv_roster$G, -gv_roster$finish)
tab1
epi.2by2(tab1, method = "cross.sectional")
rm(tab1)
```

Putting both together, we can see that both age and gender are independently associated with finisher status.

```{r}
model1 <- glm(finish ~ A + G, data = gv_roster, family = "binomial")
summary(model1)
```

The odds of finishing for females were 1/3 those of males independent of age. Since the outcome is so rare (<1%) the odds ratio closely approximates the prevalence rate ratio.

