---
title: Coronavirus in Hawaii
author: ~
date: '2020-04-04'
slug: coronavirus-in-hawaii
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---

Coronavirus is all anyone can think about these days, and the New York Times has become a repository for USA case data for some reason. They have been publishing [the data on Github](https://github.com/nytimes/covid-19-data) for now. 

```{r}
library(dplyr)
library(ggplot2)
library(scales)
covid <- readr::read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv')
```

Data pull time
```{r}
timestamp()
```

```{r}
glimpse(covid)

covid$county <- factor(covid$county)
covid$state <- factor(covid$state)
covid$fips <- as.integer(covid$fips)

covid %>% filter(state == "Hawaii") %>% glimpse
```

# Total Cases in USA as of 2020-04-04

Here is the total nationwide cases as of 2020-04-04.

```{r}
x <- covid %>% filter(date == "2020-04-04") %>% select(cases)
sum(x$cases)
```


## Total Cases by State
```{r}

covid %>% filter(date == "2020-04-04") %>%
  group_by(state) %>%
  summarize(total = sum(cases)) %>%
  arrange(-total)
```

```{r}
covid %>% filter(date == "2020-04-04") %>%
  group_by(state) %>%
  summarize(total = sum(cases)) %>% 
  filter(total > median(total)) %>% 
  arrange(-total) %>%
  ggplot(aes(x = reorder(state, total), y = total)) + 
  geom_bar(stat = 'identity') +
  scale_y_continuous(trans=log10_trans()) +
  coord_flip()

covid %>% filter(date == "2020-04-04") %>%
  group_by(state) %>%
  summarize(total = sum(cases)) %>% 
  filter(total < median(total)) %>% 
  arrange(-total) %>%
  ggplot(aes(x = reorder(state, total), y = total)) + 
  geom_bar(stat = 'identity') +
  scale_y_continuous(trans=log10_trans()) +
  coord_flip()
```

# Total Cases in Hawaii
In Hawaii, the number of cases has been growing steadily.

```{r}
covid %>% filter(state == "Hawaii", date <= "2020-04-04") %>%
  group_by(date) %>%
  summarize(total = sum(cases)) %>%
  arrange(-total)

covid %>% filter(state == "Hawaii", date <= "2020-04-04") %>%
  group_by(date) %>%
  summarize(total = sum(cases)) %>% 
  ggplot(aes(x = date, y = total)) +
  geom_point() +  
  geom_line()
```

# Cases by County in Hawaii

Honolulu county has the most cases by far (as it has by far the largest population).

```{r}
covid %>% filter(state == "Hawaii", date <= "2020-04-04") %>%
  ggplot(aes(x = date, y = cases, color = county)) +
  geom_point() +  
  geom_line()
```

# Cases by County other than Honolulu

To see the trend for the other counties, it is best to remove Honolulu County.

```{r}
covid %>% filter(state == "Hawaii" & county != "Honolulu", date <= "2020-04-04") %>%
  ggplot(aes(x = date, y = cases, color = county)) +
  geom_point() +  
  geom_line()
```

# Change per Day
Here's the change in our state per day. The growth rate in positive tests may still be accelerating. 

```{r}
covid %>% filter(state == "Hawaii", date <= "2020-04-04") %>% 
  group_by(date) %>%
  summarize(total = sum(cases)) %>% mutate(day_before_total = lag(total)) %>%
  mutate(change = total - day_before_total) %>%
  filter(!is.na(change)) %>%
  ggplot(aes(x = date, y = change)) +
  geom_point() + geom_line() + geom_smooth()
```