---
title: Last Day Distance in GVRAT
author: ~
date: '2020-06-27'
slug: last-day-distance-in-gvrat
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---

As people start finishing the race it seems like they really push themselves on the last day. I've seen people running 20, 30, 50 miles to reach the final distance. I wondered if there was a difference between what people average and what they do on the last day of the race.

# Data Loading
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
gv <- read_csv("../datasets/GVRAT_June_Summary.csv")
```

It took a little effort to reshape the data for manipulation. I wanted to get Bib, date, miles, and cumulative miles.

```{r}
gv_miles <- gv %>% 
  filter(Bib >= 1) %>% # Bibs between 0 and 1 were buzzards
  filter(Miles < 684.34) %>% # We want people who hadn't finished in May
  mutate(`31-May` = Miles) %>% # Rename Miles to a date for May
  select(-c(Name, Miles)) %>% # Drop Miles and Name
  pivot_longer(contains("-"), names_to = "run_date", 
               values_to = "miles_day") %>% # Convert to long form
  mutate(run_date = dmy(paste0(run_date, ", 2020"))) %>% 
  group_by(Bib) %>% 
  arrange(run_date) %>%
  mutate(miles_total = cumsum(miles_day)) 
```

# June Finishers

## Who finished in June?

I found people whose total miles were over the race distance and then took the first run that was over the finish distance.

```{r}
june_finishers <- gv_miles %>% 
  filter(miles_total > 684.34) %>%
  group_by(Bib) %>%
  arrange(miles_total) %>%
  slice(1)

june_finishers
```

## How many people finished each day in June so far

I constructed a graph to show how more and more people are finishing every day.

```{r}
gv_miles %>% 
  filter(miles_total > 684.34) %>%
  group_by(Bib) %>%
  arrange(miles_total) %>%
  slice(1) %>%
  group_by(run_date) %>%
  filter(run_date < ymd('2020-06-27')) %>%
  summarize(finishers = n()) %>%
  ggplot(aes(x=run_date, y = finishers)) + geom_line()
```

## June Finishers Mileage Excluding Last Day

Here's the mean miles of people who finished in June (excluding their last day).

```{r}
gv_miles %>% filter(Bib %in% june_finishers$Bib) %>%
  filter(run_date > ymd('2020-05-31')) %>%
  filter(miles_total < 684.34) %>%
  group_by(Bib) %>% 
  summarize(mean_june_miles = mean(miles_day))
```

```{r}
gv_miles %>% filter(Bib %in% june_finishers$Bib) %>%
  filter(run_date > ymd('2020-05-31')) %>%
  filter(miles_total < 684.34) %>%
  group_by() %>% 
  summarize(mean_june_miles = mean(miles_day), 
            median_june_miles = median(miles_day)) 
```

Among June finishers, the daily mileage excluding the last day was 14.8 miles.

## June Finishers Mileage on Last Day

```{r}
gv_miles %>% 
  filter(miles_total > 684.34) %>%
  group_by(Bib) %>%
  arrange(miles_total) %>%
  slice(1) %>%
  group_by() %>%
  summarize(mean(miles_day), median(miles_day), min(miles_day), max(miles_day))
```

Among the 515 finishers in June so far, the mean last day was 19.9 miles. This was skewed quite a bit by some outliers since it was so much higher than the median.

I compared the two histograms:

```{r}
gv_miles %>% filter(Bib %in% june_finishers$Bib) %>%
  filter(run_date > ymd('2020-05-31')) %>%
  filter(miles_total < 684.34) %>% 
  ggplot(aes(miles_day)) + geom_histogram(binwidth = 2) +
  ggtitle("Miles per Run before Finishing Date")

gv_miles %>% 
  filter(miles_total > 684.34) %>%
  group_by(Bib) %>%
  arrange(miles_total) %>%
  slice(1) %>%
  ggplot(aes(miles_day)) + geom_histogram(binwidth = 2) +
  ggtitle("Miles per Run on Finishing Date")
```

They are pretty similar. Note that the height of the bins is greater in the first histogram. It looks like people aren't really going crazy all out on that last day.


