---
title: Password Strength and Tidy Tuesday
author: ~
date: '2020-01-16'
slug: password-strength-and-tidy-tuesday
categories: 
  - R
tags: 
  - tidytuesday
type: ''
subtitle: ''
image: ''
---

I just found out about Tidy Tuesday, an educational exercise from the R for Data Science folks. The idea is that they publish a dataset every Tuesday for people to play around with. If you make something you're proud of you can publish it on Twitter and use the hashtag #tidytuesday. I haven't tried it before, but today's one was a topic that was of interest to me--computer security and passwords.

# Data Loading

I figured out how to load directly from the online csv.

```{r, cache = T}
library(tidyverse)
passwords <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv')
passwords <- tbl_df(passwords)

glimpse(passwords)
```

# Data Munging

The data variable types were already pretty well auto-classified but the category needed to be made into a factor. 
```{r}
passwords$category <- as.factor(passwords$category)
```


# Exploration
I figured I would look at them by category. Here is the distribution by category. Some had NA category.

```{r}
table(passwords$category, useNA = "ifany")
prop.table(table(passwords$category, useNA = "ifany"))

passwords <- passwords %>% filter(!is.na(category))
```

Here's how hard the passwords were to crack by category. They were all pretty easy to crack.

```{r}
table(passwords$offline_crack_sec)

passwords %>% ggplot(aes(x = category, y = offline_crack_sec)) + 
  geom_boxplot() + scale_y_log10() + coord_flip()

```

# Popularity vs Difficulty to Crack

Was there a correlation between how popular they were and how easy they were to crack?

```{r}
passwords %>% mutate(rank_crack = min_rank(offline_crack_sec)) %>%
  ggplot(aes(x = rank, y = rank_crack, color = category)) + 
  geom_point()
```

Not clearly.

# Conclusion

I couldn't figure out much to do with this dataset but at least I got an idea of how to load a raw csv from Github.