---
title: Coronavirus Positive Tests in Hawaii
author: ~
date: '2020-04-26'
slug: coronavirus-positive-tests-in-hawaii
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---

Today I decided to look at the tests for coronavirus that are being done in Hawaii. It looks like the number of positive tests are slowing down. I had a sense that the population is also probably healthier from social distancing, so I wanted to look at the trend in tests and positive tests.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(tidyr)
library(zoo)
```

```{r}
covid <- readr::read_csv('https://covid19-hawaii.herokuapp.com/hawaii_daily.sqlite/hawaii_daily?_format=csv&_size=max')
timestamp() # Data pull time
```


```{r}
covid$Date <- mdy(covid$Date)

covid %>% select(Date, test_total = `Total Tests`, test_daily = `Daily Tests`,
                 test_total_neg = `Negative Tests`, test_total_pos = `Positive Tests`,
                 test_total_inconcl = `Inconcl Tests`) %>%
  mutate(daily_positive = test_total_pos - lag(test_total_pos),
         daily_negative = test_total_neg - lag(test_total_neg),
         daily_inconcl = test_total_inconcl - lag(test_total_inconcl)) %>%
  select(Date, test_daily, daily_positive, daily_negative, daily_inconcl)
```

Looking at the data there seems to be a typo for the total negative tests on 4/23. It went down compared to the prior day and the positive, negative, and inconclusive tests columns add up to 1000 less than the total test column for that day. I'll just clean it for the purposes of this illustration.

```{r}
covid_edit <- covid %>% select(Date, test_total = `Total Tests`, 
                               test_daily = `Daily Tests`,
                               test_total_neg = `Negative Tests`, 
                               test_total_pos = `Positive Tests`,
                               test_total_inconcl = `Inconcl Tests`)

covid_edit$test_total_neg[which(covid_edit$Date == "2020-04-23")] <- 26017
```

Here's the curve that we often see on cumulative positive and negative tests.

```{r}
# This block uses two new functions to me, starts_with() and pivot_longer(), which is like gather but with different syntax

covid_edit %>% select(Date, test_total_neg, test_total_pos, test_total_inconcl) %>%
  pivot_longer(starts_with("test"), names_to = "test_type",
               values_to = "tests") %>%
  mutate(test_type = factor(test_type, levels = c("test_total_pos",
                                                  "test_total_inconcl",
                                                  "test_total_neg"))) %>%
  ggplot(aes(x = Date, y = tests, fill = test_type)) + geom_bar(stat = "identity")
```

It might be nice to see the positive tests per day compared to the negative tests, inconclusive, etc.

```{r}

covid_edit %>%   
  mutate(daily_positive = test_total_pos - lag(test_total_pos),
         daily_negative = test_total_neg - lag(test_total_neg),
         daily_inconcl = test_total_inconcl - lag(test_total_inconcl)) %>%
  select(Date, starts_with("daily")) %>% 
  pivot_longer(cols = starts_with("daily"), names_to = "test_type",
               values_to = "tests") %>%
  ggplot(aes(x = Date, y = tests, fill = test_type)) + geom_bar(stat = "identity")
```

Oh man, what a mess! It's impossible to see the positives because they are so small as a percentage of the total tests! A better idea might be just to plot the percentage positive anyway.

```{r}
covid_edit %>%   
  mutate(daily_positive = test_total_pos - lag(test_total_pos),
         daily_negative = test_total_neg - lag(test_total_neg),
         daily_inconcl = test_total_inconcl - lag(test_total_inconcl)) %>%
  select(Date, test_daily, daily_positive) %>%
  mutate(daily_positive_percentage = daily_positive/test_daily * 100) %>%
  ggplot(aes(x = Date, y = daily_positive_percentage)) + 
  geom_point() +
  geom_smooth(se = F) +
  ggtitle("Percent Positive Daily Coronavirus Tests, Hawaii") +
  ylab("Percent Positive")
```

That's a lot better. We can see that the percentage of positive tests is going down as the total number of cases also declines.

# Running Previous Curves
Here I updated all the other coronavirus curves for 4/26/2020.

```{r}
covid %>% filter(!is.na(`New Cases`)) %>%
  ggplot(aes(x = Date, y = `New Cases`)) + 
  geom_point() +
  geom_line() + 
  geom_smooth(se = F) +
  labs(title ="New Coronavirus Positive Tests, Hawaii",
       subtitle = "With loess smoothing line") +
  ylab("Positive Tests")
```

This loess smoothed curve is a little silly now that the cases are so low since the smoothing makes the curve dip below 0.

I updated the rolling mean curve to look a little nicer and show the difference between the day's count and the rolling mean with different shapes.

```{r}
covid %>% select(Date, `New Cases`) %>%
  mutate(rolling_mean_7d = rollmean(`New Cases`, 7, align = 'right', fill = NA)) %>% 
  mutate(newcases = `New Cases`) %>%
  select(Date, newcases, rolling_mean_7d) %>%
  ggplot(aes(x = Date)) + 
  geom_bar(aes(y = newcases), stat = 'identity', fill = 'darkolivegreen3') +
  geom_line(aes(y = rolling_mean_7d), color = 'red') +
  labs(title ="New Coronavirus Positive Tests, Hawaii",
       subtitle = "With 7 Day Rolling Mean") +
  ylab("Positive Tests")
```

```{r}
covid %>% 
  select(Date, total_cases = `Total Cases`, total_released = `Total Released`) %>%
  mutate(active_cases = total_cases - total_released) %>%
#  gather(total_cases, total_released, key = "case_type", value = "cases") %>%
  ggplot(aes(x = Date, y = active_cases)) + 
  geom_point() +
  geom_line() +
  ylim(0, 400) + xlim(ymd('2020-03-24'), ymd('2020-04-30')) +
  ggtitle("Active Cases of Coronavirus, Hawaii")
```


# License
These data are subject to the following license:

Data license: [Creative Commons Attribution 4.0 International](https://creativecommons.org/licenses/by/4.0/) · Data source: [Community Maintained Daily Hawaii COVID-19 Metrics](https://docs.google.com/spreadsheets/d/1WTUg75_mm35Z2C6PQd5AOKtVeSNfci5Mp0G_pgYTqQI/edit#gid=0) · About: This is a community maintained, unofficial table of COVID-19 stats compiled from DOH and media reports. Accuracy is not guaranteed. Please see the [Data source link]((https://docs.google.com/spreadsheets/d/1WTUg75_mm35Z2C6PQd5AOKtVeSNfci5Mp0G_pgYTqQI/edit#gid=0)) to report any errors.
