---
title: Hawaii Coronavirus Update 2020-04-14
author: ~
date: '2020-04-14'
slug: hawaii-coronavirus-update-2020-04-14
categories: []
tags: 
  - coronavirus
type: ''
subtitle: ''
image: ''
---

In my last post I used data from the [New York Times Github repository](https://github.com/nytimes/covid-19-data). The authors update it within a day or two but I found a new [crowdsourced data source](https://docs.google.com/spreadsheets/d/1WTUg75_mm35Z2C6PQd5AOKtVeSNfci5Mp0G_pgYTqQI/edit#gid=0) from the local tech community that is typically updated within minutes of the Department of Health data posting. 

# Data Loading 
The [Google Sheet](https://docs.google.com/spreadsheets/d/1WTUg75_mm35Z2C6PQd5AOKtVeSNfci5Mp0G_pgYTqQI/edit#gid=0) is very nice to look at but I've had issues recently with loading data directly from Google sheets.  Fortunately the community also made available the data via a [heroku app](https://covid19-hawaii.herokuapp.com/) that allows [CSV file download](https://covid19-hawaii.herokuapp.com/hawaii_daily.sqlite/hawaii_daily?_format=csv&_size=max).

## Packages
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(tidyr)
library(zoo)
```

## Data Download
```{r, cache = TRUE}
covid <- readr::read_csv('https://covid19-hawaii.herokuapp.com/hawaii_daily.sqlite/hawaii_daily?_format=csv&_size=max')
timestamp() # Data pull time
```


## Data processing steps.
```{r}
covid$Date <- mdy(covid$Date)
```


# Change per Day
Here's the change in our state per day, smoothed with the loess method (regression line fitting).

```{r}
covid %>% filter(!is.na(`New Cases`)) %>%
  ggplot(aes(x = Date, y = `New Cases`)) + 
  geom_point() +
  geom_line() + 
  geom_smooth() +
  labs(title ="New Coronavirus Positive Tests, Hawaii",
       subtitle = "With loess smoothing line") +
  ylab("Positive Tests")
```

I also tried smoothing by drawing a 7 day running average. I used a (new to me) function from the zoo time series analysis package. 

```{r}
covid %>% select(Date, `New Cases`) %>%
  mutate(rolling_mean_7d = rollmean(`New Cases`, 7, align = 'right', fill = NA)) %>% 
  mutate(newcases = `New Cases`) %>%
  select(Date, newcases, rolling_mean_7d) %>%
  gather(`rolling_mean_7d`, `newcases`, key = seriesname, value = cases) %>%
  ggplot(aes(x = Date, y = cases, color = seriesname)) + 
  geom_point() +
  geom_line() +
  labs(title ="New Coronavirus Positive Tests, Hawaii",
       subtitle = "With 7 Day Rolling Mean") +
  ylab("Positive Tests")
  
```

# Discussion
Both the loess smoothed and rolling 7 day mean curves are pretty nice, showing a reduction in the new positive tests per day in the last 2 weeks. There is speculation that the 5 positive tests on April 13 was due to slow test reporting from the Easter weekend. Today (4/14) had more than twice the previous day. In any case, I would guess the overall trend is downards now and I hope it continues!