---
title: COVID-19 School Indicators Update 2021-02-27
author: ~
date: '2021-02-27'
slug: udpate-covid-19-school-indicators
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---

Since the DOH does not publish data on how Hawaii fares compares to the CDC school reopening criteria, and since Maui is getting out of control, I figured I would update my graphs from last week. See that post for more context on where these numbers come from.

I again downloaded the data from the "How is Hawaii Doing at Flattening the Epidemic Curve?" Tableau figure on the [Hawaii DOH COVID-19 data site](https://health.hawaii.gov/coronavirusdisease2019/what-you-should-know/current-situation-in-hawaii/). It has daily counts by county and the positive and total tests.

```{r}
library(tidyverse)
library(lubridate)

# Read in data
dat <- read_csv("../datasets/DOH_COVID20210227.csv")
dat

# Make variables into right classes
dat$County <- factor(dat$County)
dat$Date <- mdy(dat$Date)

# County population data from 2019 US Census
county_pops <- data.frame(County = c("Hawaii", "Honolulu", "Kauai", "Maui", "Missing"), 
                          pops = c(201513, 974563, 72293, 167417, NA))

# Make case count graph
dat %>% left_join(county_pops, by = "County") %>%
  group_by(County) %>%
  arrange(Date) %>% 
  mutate(total_cases = cumsum(NewCases)) %>%
  mutate(cases_7d_ago = lag(total_cases, 7, default = 0)) %>%
  mutate(cases_7d_sum_per100k = (total_cases-cases_7d_ago)/pops*100000) %>%
  filter(Date > "2020-12-31") %>%
  ggplot(aes(x = Date, y = cases_7d_sum_per100k, color = County)) +
  geom_line() +
  labs(title = "Total Cases Per 100,000 in Last 7 Days", 
       subtitle = "CDC Indicators to Inform Decision Making") +
  ylab("Total cases over last 7 days per 100,000 population") +
  geom_hline(yintercept = 9.5, linetype = "dotted", color = "blue") +
  geom_hline(yintercept = 49.5, linetype = "dotted", color = "yellow") +
  geom_hline(yintercept = 99.5, linetype = "dotted", color = "orange")

# Make Test Positive Graph
dat %>% 
  group_by(County) %>%
  arrange(Date) %>% 
  mutate(total_tests = cumsum(TotalTestEncounters)) %>%
  mutate(total_positives = cumsum(NewPositiveTests)) %>%
  mutate(total_tests_7d_ago = lag(total_tests, 7, default = 0)) %>%
  mutate(total_positives_7d_ago = lag(total_positives, 7, default = 0)) %>%
  mutate(test_positive_rate_7d = (total_positives-total_positives_7d_ago) /
           (total_tests - total_tests_7d_ago)) %>%
  select(County, Date, test_positive_rate_7d) %>%
  filter(Date > "2020-12-31") %>%
  ggplot(aes(x = Date, y = test_positive_rate_7d*100, color = County)) +
  geom_line() +
  labs(title = "Test Positive Rate in Last 7 Days", 
       subtitle = "CDC Indicators to Inform Decision Making") +
  ylab("Test positive rate over last 7 days (%) ") +
  geom_hline(yintercept = 5, linetype = "dotted", color = "blue") +
  geom_hline(yintercept = 7.95, linetype = "dotted", color = "yellow") +
  geom_hline(yintercept = 9.95, linetype = "dotted", color = "orange")
```

Oops, as I was writing this I discovered I had written a helper function to help with the calculations...

```{r}
rollsum <- function(x, lag_start, lag_end) {
  require(dplyr)
  total = 0
  for(i in lag_start:lag_end) {
    total <- total + lag(x, i, default = 0)
  }
  return(total)
}
```

Time to rewrite the code for those plots using the new function! It saves some lines of code but also seems to make it more readable to me.

```{r}
# Make case count graph
dat %>% left_join(county_pops, by = "County") %>%
  group_by(County) %>%
  arrange(Date) %>% 
  mutate(cases_7d_sum_per100k = rollsum(NewCases, 0, 6)/pops*100000) %>%
  filter(Date > "2020-12-31") %>%
  ggplot(aes(x = Date, y = cases_7d_sum_per100k, color = County)) +
  geom_line() +
  labs(title = "Total Cases Per 100,000 in Last 7 Days", 
       subtitle = "CDC Indicators to Inform Decision Making") +
  ylab("Total cases over last 7 days per 100,000 population") +
  geom_hline(yintercept = 9.5, linetype = "dotted", color = "blue") +
  geom_hline(yintercept = 49.5, linetype = "dotted", color = "yellow") +
  geom_hline(yintercept = 99.5, linetype = "dotted", color = "orange")

# Make Test Positive Graph
dat %>% 
  group_by(County) %>%
  arrange(Date) %>% 
  mutate(test_positive_rate_7d = rollsum(NewPositiveTests, 0, 6) / 
           rollsum(TotalTestEncounters, 0, 6)) %>%
  select(County, Date, test_positive_rate_7d) %>%
  filter(Date > "2020-12-31") %>%
  ggplot(aes(x = Date, y = test_positive_rate_7d*100, color = County)) +
  geom_line() +
  labs(title = "Test Positive Rate in Last 7 Days", 
       subtitle = "CDC Indicators to Inform Decision Making") +
  ylab("Test positive rate over last 7 days (%) ") +
  geom_hline(yintercept = 5, linetype = "dotted", color = "blue") +
  geom_hline(yintercept = 7.95, linetype = "dotted", color = "yellow") +
  geom_hline(yintercept = 9.95, linetype = "dotted", color = "orange")

```