---
title: Advocacy Survey
author: ~
date: '2021-05-22'
slug: advocacy-survey
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---
(Updated 2021-09-01) 
Our pediatric chapter's advocacy committee asked the chapter members what their priorities are for advocacy issues. We gave them a list of topics to rate as 1-10 and I needed to calculate the mean score then rank them. This is probably doable as a pivot table within the spreadsheet software, but why pivot table when I can practice some data manipulation skills?

First load tidyverse package and advocacy dataset.
```{r}
library(tidyverse)
advo <- read_csv("../datasets/advocacy_survey.csv")
advo
```

I got rid of the timestamp and the free text questions.
```{r}
advo <- advo[-c(1,20,21)]
```

Then I had to convert the wide data into long data, calculate the mean responses, and sort them by mean.
```{r}
advo %>%  
  pivot_longer(cols = names(advo),
               names_to = "topic", 
               values_to = "score") %>%
  group_by(topic) %>%
  summarize(responses = sum(!is.na(score)), mscore = mean(score, na.rm = T)) %>% 
  arrange(-mscore)
```



